---
title: "lab2"
author: "tommaso tarchi"
date: "2022-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### CLT

We compute the probability that Posillipo win the match, by using the Skellam distribution:
```{r}
library(skellam)
lambda <- 5   # parameters of the Poisson distributions
mu <- 7

PWin_P_sk <- pskellam(0, lambda, mu, lower.tail = FALSE)
PWin_P_sk
```

We now use the normal approximation for Poisson distributions, using the fact that a Poisson with rate $\lambda$ can be approximated by a normal with mean and variance both equal to $\lambda$:
```{r}
lambda <- 5
mu <- 7

# we call z the variable x-y (with x scores by Posillipo and y scores by the other team)
# the two values below are normal approximations
mean.z <- lambda - mu
sd.z <- sqrt(lambda + mu)

PWin_P_n <- pnorm(0, mean.z, sd.z, lower.tail = FALSE)
PWin_P_n
```
```{r}
# using continuity correction
PWin_P_n_cor <- pnorm(0.5, mean.z, sd.z, lower.tail = FALSE)
PWin_P_n_cor
```
```{r}
# plotting
plot(function(x) dnorm(x, mean.z, sd.z), xlim = c(-15, 25), ylim = c(0, 0.3))
points(0:25, dpois(0:25, lambda), col = "blue")
points(0:25, dpois(0:25, mu), col = "red")
points(-15:25, dskellam(-15:25, lambda, mu), col = "green")
```

### Distribution of the sample mean and variance

We perform a simulation with R samples of size n:
```{r}
set.seed(1234)
R <- 1000
n <- 30

# generate the sample of size n (from the distribution aforementioned) R times 
samples <- array(NA, c(3, R, n))
for (i in 1:R) {
  samples[1, i, ] <- rnorm(n, 0, 1)
  samples[2, i, ] <- rt(n, df = 3)
  samples[3, i, ] <- runif(n, 0, 1)
}

# compute the sample mean and the sample variance
# (apply() la funzione sulle righe (parametro=1) o sulle colonne (par=2) della matrice data)
samples_stat <- array(NA, c(3, 2, R))
for (i in 1:3) {
  samples_stat[i, 1, ] <- apply(samples[i, , ], 1, mean)
  samples_stat[i, 2, ] <- apply(samples[i, , ], 1, var)
}

# visualize the results
par (mfrow=c(1,3))

hist(samples_stat[1, 1, ], nclass = 30, probability = TRUE, 
     xlab="y", main= "N(0,1)", cex.main = 1.5)
curve(dnorm(x, 0, sqrt(1/n)), add = TRUE, col = "red", lwd = 2)

hist(samples_stat[2, 1, ], nclass = 30, probability = TRUE, 
     xlab = "y", main = expression(t[3]), cex.main = 1.5)
curve(dnorm(x, 0, sqrt(  (3/((3 - 2) * n)))), add = TRUE, col = "red", lwd = 2)

hist(samples_stat[3, 1, ], nclass = 30, probability = TRUE, 
     xlab = "y", main = "U(0,1)", cex.main = 1.5)
curve(dnorm(x, 1/2, sqrt(1/(12 * n))), add = TRUE, col = "red", lwd = 2)
```

It is known that the sample variance $S^2$ of the mean of a normal distribution is distributed proportionally to a $\chi^2$ distribution with $n-1$ degrees of freedom
$$
\frac{(n-1)S^2}{\sigma^2} \sim \chi^2_{n-1}
$$
```{r}
# visualizing the sample variance for standard normal distrbution
par (mfrow=c(1,1))
sigma <- 1 
hist(samples_stat[1, 2, ], nclass = 30, probability = TRUE, 
     xlab=expression(s^2), main = "", cex.main = 1.5)
curve(((n - 1)/sigma^2) * dchisq(x * ((n - 1)/sigma^2), df = n - 1),
      add = TRUE, col = "red", lwd = 2)
```

### biased and unbiased estimators for the variance

We could also define a biased estimator for the variance
$$
S_b := \frac{1}{n} \sum_{i=1}^{n}\left(x_i-\mu\right)^2
$$
We now confront biased and unbiased estimators for the variance of a normal distribution:
```{r}
set.seed(2)
R <- 1000
n <- 10
mu <- 0  # real expectation value
sigma <- 1  # real standard deviation

s2 <- rep(NA, R)  # for unbiased estimator
s2_b <- rep(NA, R)  # for biased estimator

# (la funzione var() usa l'unbiased estimator)
for ( i in 1 :R ) {
  y <- rnorm(n, mu, sigma)
  s2[i] <- var(y) 
  s2_b[i] <- var(y)*(n-1)/n
}

s2_mean <- mean(s2)
s2_b_mean <- mean(s2_b)

# plotting
# (in red the true mean, in blue the estimated mean)
par(mfrow=c(1,2), oma=c(0,0,0,0))
hist(s2, breaks = 50, xlab = expression(s^2), probability = TRUE,
     main = expression(s^2), cex.main = 1.5)
abline(v = sigma^2, col = "red", lwd = 2)
abline(v = s2_mean, col = "blue", lwd = 2, lty = 2) 

hist(s2_b, breaks = 50, xlab = expression(s[b]^2), probability = TRUE,
     main = expression(s[b]^2), cex.main = 1.5)
abline(v = sigma^2, col = "red", lwd = 2)
abline(v = s2_b_mean, col = "blue", lwd = 2, lty = 2) 
```

### mean estimators for normal distribution

In addition to sample mean and median we can define for a normal distribution the following mean estimators:

1) semi-sum of minimum an maximum:
$$
ssum(x) := \frac{max(x)+min(x)}{2}
$$
2) the 10% trimmed sample mean: sample mean computed discarding 10% upper and 10% lower data

In the following we will call $\hat\mu_1$ the sample mean, $\hat\mu_2$ the sample median, $\hat\mu_3$ the semi-sum of the extremes and $\hat\mu_4$ the trimmed sample mean.
```{r}
R <- 1000
n <- 10
mu <- 2
sigma <- 2

est <- matrix(0, R, 4)
means.means <- rep(NA, 4)   # mean values of estimators

set.seed(1234)
for (i in 1:R) {
 x <- rnorm(n, mu, sigma)
 est[i, 1] <- mean(x)
 est[i, 2] <- median(x)
 est[i, 3] <- (min(x) + max(x))/2
 est[i, 4] <- mean(x, trim = 0.1)
}

means.means <- apply(est, 2, mean)

par(mfrow = c(1, 1), xaxt = "n")
boxplot(est, main="Comparison between four estimators")
par(xaxt = "s")
axis(1, 1:4, c(expression(hat(mu)[1]), expression(hat(mu)[2]), expression(hat(mu)[3]), expression(hat(mu)[4])) )
abline(h = mu, lwd = 2, col = "blue")
```
We now compute the bias:
```{r}
bias <- apply(est, 2, mean) - mu
bias
```
```{r}
variance <- apply(est, 2, var)
variance
```
All the estimators appear unbiased and the sample mean register the lowest estimated sample variance, which is a good approximation of $σ_2/n=0.4$.

To check the consistency of the estimators we have to increase the size of the samples:
```{r}
n2 <- 200

est2 <- matrix(0, R, 4)

set.seed(1234)
for (i in 1:R) {
 x <- rnorm(n2, mu, sigma)
 est2[i, 1] <- mean(x)
 est2[i, 2] <- median(x)
 est2[i, 3] <- (min(x) + max(x))/2
 est2[i, 4] <- mean(x, trim = 0.1)
}

means.means2 <- apply(est2, 2, mean)
```
```{r}
# confronting sample means
par(mfrow = c(1, 2))
hist(est[, 1], probability = TRUE, main = "n=10")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means[1], col = "blue", lwd = 2, lty = 2)
hist(est2[, 1], probability = TRUE, xlim = c(0, 4), main = "n=200")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means2[1], col = "blue", lwd = 2, lty = 2)
```
```{r}
# confronting sample medians
par(mfrow = c(1, 2))
hist(est[, 2], probability = TRUE, main = "n=10")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means[2], col = "blue", lwd = 2, lty = 2)
hist(est2[, 2], probability = TRUE, xlim = c(-1, 5), main = "n=200")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means2[2], col = "blue", lwd = 2, lty = 2)
```
```{r}
# confronting semisum of the extremes
par(mfrow = c(1, 2))
hist(est[, 3], probability = TRUE, main = "n=10")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means[3], col = "blue", lwd = 2, lty = 2)
hist(est2[, 3], probability = TRUE, xlim = c(-1, 5), main = "n=200")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means2[3], col = "blue", lwd = 2, lty = 2) 
```
```{r}
# confronting 10% trimmed mean
par(mfrow = c(1, 2))
hist(est[, 4], probability = TRUE, main = "n=10")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means[4], col = "blue", lwd = 2, lty = 2) 
hist(est2[, 4], probability = TRUE, xlim = c(0, 4), main = "n=200")
abline(v = mu, col = "red", lwd = 2)
abline(v = means.means2[4], col = "blue", lwd = 2, lty = 2) 
```
All of the four estimators seem to be consistent, since their variance lowers as the samples get larger.